### 2018-04-05 dissertation/2-complexnetworks/chapter2.tex
Extended chapter 5 and small improvements elsewhere

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_1.png}
\caption{Initial topology and trust graph} \label{fig:2_2a}
\end{subfigure}

\vspace*{\fill} % separation between the subfigures

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_2.png}
\caption{Final topology graph} \label{fig:2_2b}
\end{subfigure}

\vspace*{\fill} % separation between the subfigures

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_trust.png}
\caption{Final trust graph} \label{fig:2_3b}
\end{subfigure}

>>>>

\autoref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
The topology graph changes as one node moves across the network, with edges being added and removed as nodes move into and out of each other's ranges.
The trust graph, however, is constantly being built, maintaining past relationships even if the nodes are no longer in communication range.

@@

\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_1.png}
\caption{Initial topology graph} \label{fig:2_2a}
\end{subfigure}

\vspace*{0.5cm} % separation between the subfigures

\begin{subfigure}{0.08\textwidth}
\includegraphics[width=\linewidth]{images/manet_trust_1.png}
\caption{} \label{fig:2_3b}
\end{subfigure}

\vspace*{0.5cm} % separation between the subfigures

\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_2.png}
\caption{Final topology graph} \label{fig:2_2c}
\end{subfigure}

\vspace*{0.5cm} % separation between the subfigures

\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\linewidth]{images/manet_trust_2.png}
\caption{Final trust graph} \label{fig:2_3d}
\end{subfigure}

### 2018-04-05 dissertation/4-project/chapter4.tex
Extended chapter 5 and small improvements elsewhere

@@

This is necessary because, in the event that a node converts from benign to malicious, old information might still indicate that it is benign, as shown in ------figure------.

Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
This is done by setting a maximum age for edges.
Once a edge is older than that maximum value, it is discarded.
A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).

>>>>

\begin{subfigure}{0.30\textwidth}
\includegraphics[width=\linewidth]{images/rogue/0.png}
\caption{All nodes are benign and belong to the same strongly connected component.} \label{fig:rogue0}
\end{subfigure}
\hspace*{1cm} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/rogue/1.png}
\caption{Node $0$ becomes malicious, but still belongs to the component.} \label{fig:rogue1}
\end{subfigure}

\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/rogue/2.png}
\caption{Node $2$ updates its opinion of node $0$, identifying it as malicious.} \label{fig:rogue2}
\end{subfigure}
\hspace*{1cm} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/rogue/3.png}
\caption{$3$'s opinion becomes too old and is discarded.} \label{fig:rogue3}
\end{subfigure}

\caption{Example of what happens when a node becomes malicious.}\label{fig:rogue}
\end{figure}

@@

Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
This is done by setting a maximum age $m$ for edges; once a edge is older than $m$, it is discarded.
A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).

Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \autoref{fig:rogue}.
Immediately after node $0$ becomes malicious, it is still part of a benign strongly connected component, because there are paths from every other node to $0$, and from $0$ to every other node.
Some time later, node $2$ interacts with $0$ again and identifies it as malicious; other nodes that interact with $2$ will receive that updated information.
However, the outdated edge $3\rightarrow 0$ keeps $0$ in the component, since other nodes trust $3$ and $3$ still hasn't updated its own opinion of $0$.
Once that edge is old enough, it is discarded and node $0$ is correctly identified as malicious by all other nodes.

### 2018-04-05 dissertation/5-simulations/chapter5.tex
Extended chapter 5 and small improvements elsewhere

\autoref{fig:randommalicious} shows the variation of results for different amounts of malicious nodes in the network.
By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network, although there is still a small amount of false positive detections.
At 40\%, a small part of malicious nodes are yet to be detected and the standard deviation is larger overall.
At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is considered completely compromised.
True positive and false positive detections are often inverted, because whenever there are more malicious nodes than benign one in a certain node's network model, the algorithm will classify the malicious ones as correct.
The amount of malicious nodes also affects how quickly nodes are able to acquire information about the network, since nodes do not trust information from malicious neighbors.

@@

@@

\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
\caption{1\% malicious.}
\end{subfigure}
\hspace*{0.1cm} % separation between the subfigures
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
\caption{5\% malicious.}
\end{subfigure}

\vspace{1cm}

\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
\caption{10\% malicious.}
\end{subfigure}
\hspace*{0.1cm} % separation between the subfigures
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
\caption{30\% malicious.}
\end{subfigure}

\vspace{1cm}

\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
\caption{40\% malicious.}
\end{subfigure}
\hspace*{0.1cm} % separation between the subfigures
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
\caption{50\% malicious.}
\end{subfigure}

\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes.}
\label{fig:randommalicious}
\end{figure}

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
\caption{$h = 0.3$.} \label{fig:tarjan0}
\end{subfigure}

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
\caption{$h = 0.5$.} \label{fig:tarjan0}
\end{subfigure}

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
\caption{$h = 0.7$.} \label{fig:tarjan0}
\end{subfigure}

@@

>>>>

\autoref{fig:randommalicious1} and \autoref{fig:randommalicious2} show the variation of results for different amounts of malicious nodes in the network.
By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network, although there is still a small amount of false positive detections.
At 40\%, a small part of malicious nodes are yet to be detected and the standard deviation is larger overall.
At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is considered completely compromised.
True positive and false positive detections are often inverted, because whenever there are more malicious nodes than benign one in a certain node's network model, the algorithm will classify the malicious ones as correct.
The amount of malicious nodes also affects how quickly nodes are able to acquire information about the network, since nodes do not trust information from malicious neighbors.

@@

\autoref{fig:extramaliciousaging1} and \autoref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
These were the only simulations to consider information aging.
Here, several different values for the maximum age, $m$, were tested.
For the simulation, information timestamps are set as the iteration in which they were added to the graph ($s \gets i$), and the age is calculated as the difference between the current iteration and the timestamp ($age \gets i - s$).
With $m=10$, information is discarded too quickly, resulting in highly imprecise results.
Starting with $m=250$, results are reasonable, and the additional malicious node is detected approximately 250 iterations after it converts.
With higher values such as $m=1000$, detection takes longer, but not much else changes. 
When the value is very high, like in the case of $m=5000$, the extra malicious node is not detected by the end of one simulated day.

\textit{1. Decentralized trust establishment}:
TruMan is built from the ground up for decentralized systems.
Nodes form their own abstractions of the network and the model does not rely on a central observer.

\textit{2. Coping with sparsity}:
The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
Since nodes carry information from previous interactions, the model can also work on temporarily isolated parts of the network.

\textit{3. Event/task and location/time dynamics}:
Aside from the time-related information used for information aging, dynamics were not used in the simulations presented here.
However, TruMan can be extended to consider event, task, location and time dynamics when managing trust.

\textit{4. Scalability}: 
Due to the low complexity of the algorithms used in the model, TruMan is highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
It has also been demonstrated that iterations of the algorithm do not need to run very frequently in order to detect malicious nodes with high accuracy --- once every ten seconds is more than enough.

\textit{5. Integrated confidence measure}:
Since nodes using TruMan store trust values as a number between 0 and 1 that increase or decrease the more nodes interact (i.e. the more evidence they gather), this value can be used as a confidence measure of the opinion.
Setting the threshold $h$ to a value closer to 1 makes it unlikely that untrustworthy nodes will be incorrectly labeled as correct.

\textit{6. System level security}:
A public-private key solution can be used to verify message integrity.
This has not been considered in evaluations of TruMan.
However, it can be included as a separate security model during the transmission of messages.

\textit{7. Sensitivity to privacy concerns}:
TruMan has not been designed with this in mind, as it requires that nodes maintain a constant identity the whole time.
However, this does not inhibit other types of privacy protection, which can be implemented in addition to TruMan.

\textit{8. Robustness}: the model's resistance to attacks.
TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
Experiments show that, when fewer than 50\% of nodes in the network are malicious, TruMan performs as expected.
Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
Furthermore, since nodes take into consideration experiences from several trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
Experiments with information aging show that it is possible to detect nodes that suddenly become malicious, although further experimentation with known attacks is still necessary.
%Furthermore, the analysis also takes into account the complexity of the algorithm necessary to implement the trust model, because a high priority of TruMan is the

Finally, it is worth noting that, considering the scale of the problem, TruMan's cost is very low without sacrificing completeness and correctness.
The model satisfies or permits most of the desired properties of a trust model with low computational complexity, making it viable for real-world use.

@@

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
\caption{1\% malicious.}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
\caption{5\% malicious.}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
\caption{10\% malicious.}
\end{subfigure}

\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1/2).}
\label{fig:randommalicious1}
\end{figure}

@@

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
\caption{30\% malicious.}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
\caption{40\% malicious.}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
\caption{50\% malicious.}
\end{subfigure}

\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (2/2).}
\label{fig:randommalicious2}
\end{figure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
\caption{$h = 0.3$.} \label{fig:threshold03}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
\caption{$h = 0.5$.} \label{fig:threshold05}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
\caption{$h = 0.7$.} \label{fig:threshold07}
\end{subfigure}

@@

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
\caption{$m = 10$.} \label{fig:extra10}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
\caption{$m = 250$.} \label{fig:extra250}
\end{subfigure}

\caption{Simulation with information aging, with different maximum age values (1/2).}
\label{fig:extramaliciousaging1}
\end{figure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
\caption{$m = 1000$.} \label{fig:extra1000}
\end{subfigure}

\begin{subfigure}{0.6\textwidth}
\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
\caption{$m = 5000$.} \label{fig:extra5000}
\end{subfigure}

\caption{Simulation with information aging, with different maximum age values (2/2).}
\label{fig:extramaliciousaging2}
\end{figure}

### 2018-04-05 dissertation/6-conclusao/chapter6.tex
Extended chapter 5 and small improvements elsewhere

The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
By placing nodes into strongly connected components, a network containing a large amount of node can be simplified into a much smaller one.
Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
This allows for a low complexity approach to malicious node identification in a dynamic network.

>>>>

The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previously proposed algorithms in order to generate fast and accurate results.

As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
By placing nodes into strongly connected components, a network containing a large amount of nodes can be simplified into a much smaller one.
Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
This allows for a low complexity approach to malicious node identification in a dynamic network.
